---
title: "Assigning points to polygons"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{assign_points}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

There are three functions that can assign points (coordinates) to polygons:

  + **assign_points_terra** is useful when each point could overlap with only polygon in a shapefile
  + **assign_points_secr** is useful when each point could overlap with one or more polygons in a shapefile. It returns a presence absence data frame to indicate whether a point is inside a polygon. The last column in the data frame indicates all the polygons associated with each point.
  + **assign_points_to_nearest_polygon** is useful when there are points that don't overlap with any polygons, but you want to know which polygon a point is nearest to. This function uses the edges, rather than the centroids, of polygons.

The examples below assign fishing coordinates to DFO September research vessel survey strata and/or Marine Protected Areas.


```{r setup, warning = FALSE, message = FALSE,error=FALSE}
library(gslSpatial)
library(ggplot2)
library(tidyterra)
```

## Get a shapefile
Notice that the coordinate reference system (crs) is lon/lat NAD83 (EPSG:4269). When working with spatial data, your data should all have the same crs. If they don't, you can re-project them (e.g., ?terra::project).
```{r }
rv<-get_shapefile('rv.sgsl')

# see information about the object
rv 

# plot it
ggplot()+
  geom_spatvector(data=rv,aes(fill=factor(id)))
```


## Get some data
This example uses commercial landings data that have longitude and latitude coordinates with the same crs as the shapefile.
```{r results = "hold"}
df<-dat.ziff[,c('longitude','latitude')]
head(df)
```

## Plot everything
There are many points inside the polygons as well as outside the polygons
```{r results = "hold"}

ggplot()+
  geom_spatvector(data=rv)+
  geom_point(data=df,aes(longitude,latitude))+
  theme_bw()

```


## Using function `assign_points_terra`
```{r }
x<-assign_points_terra(df$longitude, df$latitude,rv)
head(x)

stratum<-x[,3]

# combine
df1<-cbind(df,stratum)
head(df1)

ggplot()+
  geom_spatvector(data=rv)+
  geom_point(data=df1,aes(longitude,latitude,col=factor(stratum)))+
  theme_bw()

```

## What if there are overlapping polyons? Use function `assign_points_secr`
If there are overlapping polygons causing your data points to 'belong' to multiple polygons, you can use function `assign_points_secr`

Start by getting another shapefile. This example uses Oceans Act Marine Protected Areas
```{r, warning = FALSE, message = FALSE }
mpa<-get_shapefile('mpa')

mpa

ggplot()+
  geom_spatvector(data=mpa,aes(fill=factor(NAME_E)),col='blue',lwd=1.25)+
  geom_spatvector(data=rv,fill=NA,col='red')+
  geom_point(data=dat.ziff,aes(longitude,latitude),alpha=0.5)+
  theme_bw()+
  theme(legend.position="bottom",legend.title=element_blank())+
  guides(fill=guide_legend(ncol=1,byrow=TRUE))
```

In order to demonstrate function `assign_points_secr`, we will combine the two shapefiles as if they were originally a single shapefile. 
```{r, warning = FALSE, message = FALSE }
library('terra')

# first need to make names match
names(rv)
names(mpa)

rv$NAME<-as.character(rv$id)
mpa$NAME<-as.character(mpa$NAME_E)

shape<-rbind(rv,mpa)
shape<-shape[, c("NAME")]
shape
```

```{r, fig.height=7, warning = FALSE, message = FALSE }
ggplot()+
  geom_spatvector(data=shape,aes(fill=factor(NAME)),alpha=0.5)+
  theme_bw()+
  theme(legend.position="bottom",legend.title=element_blank())+
  guides(fill=guide_legend(nrow=8))
```

Now use function `assign_points_secr`
```{r }
x<-assign_points_secr(dat.ziff[,'longitude'],
                        dat.ziff[,'latitude'],
                        shape,"NAME")

polygon<-x$assigned.polygon

df2<-cbind(dat.ziff,polygon)
```

Using function `table` shows that 69 points fall within the area in which stratum 416 overlaps with Banc-des-Américains Marine Protected Area. All other assigned points belong to non-overlapping polygon areas.
```{r fig.height=7}
rbind(table(df2$polygon))

ggplot()+
  geom_spatvector(data=shape,fill=NA)+
  geom_point(data=df2,aes(longitude,latitude,col=polygon),alpha=0.5)+
  theme_bw()+
  theme(legend.position="bottom",legend.title=element_blank())+
  guides(col=guide_legend(ncol=3))
```


## Assigning points to the nearest polygon
Finally, the previous examples show that a number of data points don't overlap with any of the polygons in our shapefiles. But what if we wanted to find out which polygons they were closest to? Then we could use function `assign_points_to_nearest_polygon`. This function takes a bit longer to run compared to `assign_points_terra` or `assign_points_secr`.


```{r }
# get the unassigned data points
pts.outside<-df2[which(is.na(df2$polygon)),]

x<-assign_points_to_nearest_polygon(pts.outside$longitude, pts.outside$latitude, shape, 'NAME')

head(x)

ggplot()+
  geom_spatvector(data=shape,fill=NA)+
  geom_point(data=x,aes(x,y,col=NAME),alpha=0.5)+
  theme_bw()+
  theme(legend.position="bottom",legend.title=element_blank())+
  guides(col=guide_legend(ncol=6))

```
