---
title: "Working with raster objects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{working_with_rasters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


<span style="color:red">**vignette under construction**</span>

# Objectives

  1. Make square 10 km X 10 km grid cells over the GSL
  2. Summarise fishery catch data
  3. Export the results as either a data frame or a SpatRaster
  
  
```{r setup, warning = FALSE, message = FALSE}
library(gslSpatial)
library(ggplot2)
library(terra)
library(tidyterra)
```

# Make square grids
Use the `make_grid` function, specifying the grid cell size (resolution). Using a resolution of 10 makes each grid cell 10 km X 10 km. The finer the resolution, the longer it will take to generate. By default, this function generates grid cells spanning all of NAFO 3Pn4RST, and assigns the corresponding NAFO division or subdivision to each cell.

```{r }
grid<-make_grid(10)
```

The resulting grid is a data frame. The coordinates are the centers of the grid cells. Longitude and latitude coordinates are NAD83, EPSG:4269, in decimal degrees. X and Y coordinates are NAD83, UTM zone 20N, with units in km.
```{r }
class(grid)
head(grid)
```

## Plot the grid, showing the NAFO boundaries.
Notice that grid cells without assigned NAFO divisions have centers that fall outside the NAFO borders.
```{r }
# get NAFO boundaries
naf<-get_shapefile('nafo.clipped')
naf<-terra::project(naf, '+proj=utm +zone=20 +datum=NAD83 +units=km +no_defs')

ggplot()+
  geom_tile(data=grid,aes(X,Y,fill=nafo.assigned))+
  geom_spatvector(data=naf,fill=NA)
```

## Restrict by location and depth

  + Restrict to NAFO 4T
  + Use function `get_depth` to estimate depth in meters. Use ?gslSpatial::get_depth for details.
  + Remove grid cells with depths shallower than 5 meters.
```{r }
grid<-grid[which(grid$nafo.assigned=="4T"),]

depth<-get_depth(grid$longitude,grid$latitude,"epsg:4269")
grid$depth<-unlist(depth)
str(grid)

grid<-grid[-which(grid$depth>(-5)),]

ggplot()+
  geom_tile(data=grid,aes(X,Y,fill=depth))+
  geom_spatvector(data=naf,fill=NA)+
  scale_fill_viridis_c()
```

# Get catch data

## Plot raw data

# Summarise

## Plot



